---
title: 'Lab 3: Data Wrangling on Soccer Tournament Data'
date: 'July 9, 2020'
output:
  pdf_document: default
  html_document: default
---
```{r setup, message = F, include=FALSE}
library(tidyverse)
```

Soccer tournament data wrangling
--------------------------
Read the dataset of football games.
```{r}
d <- read_csv("data/results.csv")
d
```

1. Select variables `date`, `home_team` and `away_team`.
```{r}
select(d,date,home_team,away_team)
```
2. Subset games with **Brazil** as the home team.
```{r}
d %>% filter(home_team=="Brazil")
```

3. Choose the games that Brazil won as the home team, and select variables `date`, `away_team` and `tournament`.
```{r}
d%>%filter(home_team=="Brazil" & home_score>away_score) %>% select(date,away_team,tournament)
```

4. Add the difference of goals, and an indicator variable called `goleada` for when the difference of goals is large,  and select what we did only for Brazil. **Hint: use `ifelse`**.
```{r}
Func<-function(x){
  y=vector(length=length(x))
  for (n in 1:length(x)){
    if (x[n]>3)
      y[n]=TRUE
    else
      y[n]=FALSE
  }
  c(y)
}
d1<- d %>% mutate(difference=abs(home_score-away_score)) %>% mutate(g=Func(difference))

filter(d1,home_team=="Brazil" |away_team=="Brazil")
```

5. What was the largest difference in goals within these games?
```{r}
max(d1[,"difference"])
```

6. The top 5 goleadas?
```{r}

```
7. Summary on goals scored by home teams, such as `mean` of `home_score` and `away_score`, `std`, using `group_by` and `summarise`
```{r}
d %>% group_by(home_team) %>%
  summarise(Mean_Home_Score=mean(home_score),M_A_S=mean(away_score))
```

8. Proportion of victories of **Brazil** on different tournaments against each opponent, for instance, **Argentina**.
```{r}
rm(Brazil)
Brazil<- filter(d,home_team=="Brazil" | away_team=="Brazil")
Brazil<- Brazil %>% mutate(Win=(home_team=="Brazil" & home_score>away_score) | 
                    (away_team=="Brazile" & away_score>home_score))%>% select(date,home_team,away_team,home_score,away_score,Win)

Brazil_Argen<- filter(Brazil,(home_team=="Argentina" & away_team=="Brazil")| (away_team=="Argentina"& home_team=="Brazil"))

colSums(Brazil_Argen[,"Win"])/nrow(Brazil_Argen)
```
